CorReg: model-based pretreatment for regression with correlated variables. Application in steel industry

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@480 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@481 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we don't have a system of regressions but one regression on our response variable and a system of sub-regressions (without the response variable). Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss. So we may obtain two kinds of zeros in the marginal model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.

@@

\section{Model to select decorrelated covariates}
\subsection{The decorrelation problem in regression}\label{sectionOLS}
For a model defined by 
	\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
	we have the following Ordinary Least Squares (\textsc{OLS}) estimators \cite{saporta2006probabilites}:
	\begin{equation}
		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
	\end{equation}
	With variance 
	\begin{equation}
		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1} \label{eqOLS}
	\end{equation}
	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
	\begin{eqnarray}
		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
	\end{eqnarray}	 
	and when correlations between covariates are strong, the matrix to invert is ill-conditioned, the variance explodes and so does the MSE.
	Another problem is that matricial inversion requires $n\geq p$. When it is not the case, a frequently used alternative is the Moore-Penrose generalized inverse \cite{katsikis2008fast}. In the followings, matrices inverses when $n<p$ will refer to this generalized inverse. Thus OLS can obtain some results even with $n<p$ (see section \ref{sectionsimul} ).
\subsection{Modelisation of the correlations}
Let now $X \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
We propose to explicitly define a family of $p_2$ internal regressions between covariates whith $I_2$ the set of indices of endogenous variables in $X$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$. Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $p_1=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression. Thus we have $p_2=| I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.

@@

We define a family of $p_2$ regressions within correlated covariates in $X$:
	\begin{equation}
		\forall j \in I_2: X^j_{|X_1,S}=X_1B^j_1+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n) \label{SR}
	\end{equation}
		where $B_1^j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
It remains $p-p_2$ independent exogenous variables : $\forall j \notin I_2 : X^j_{|S} \sim f(\theta_j)$

@@

This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian (see the appendices). These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $B_1^j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
\\
If there are exact regressions ($\sigma^2_j=0$) in (\ref{SR}), the structure won't be identifiable. But it only means that several structure will have the same likelihood and they will have the same interpretation. So it's not really a problem. Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete. %\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.

@@

\bibliography{biblio}{}
\bibliographystyle{plain}
\section{Appendices}
	\subsection{Identifiability of the structure}
	The model presented above relies on a discrete structure $S$ between the covariates. But to find it we need identifiability property to insure the MCMC will asymptotically find the true model. Identifiability of the discrete structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions ? The answer is no. Thus $S$ is identifiable.

>>>>

	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we have one regression on our response variable and a system of sub-regressions (without the response variable). Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss. So we may obtain two kinds of zeros in the marginal model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.

@@

\section{Model to select decorrelated covariates}
\subsection{The decorrelation problem in regression}\label{sectionOLS}
For a model defined by 
	\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ with $I_n$ the $n$-sized identity matrix,
	we have the following Ordinary Least Squares (\textsc{OLS}) estimator:
	\begin{equation}
		\hat{\beta}=\left(X'X \right) ^{-1}X'Y
	\end{equation}
	With variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\beta})=\sigma_Y^2(X'X)^{-1} \label{eqOLS}
	\end{equation}
	 The Mean Squared Error (\textsc{MSE}) on $\hat{\beta}$ is:
	\begin{eqnarray}
		\textsc{MSE}(\hat{\beta}|\beta,X)&=&\operatorname{Bias}^2+\operatorname{Tr}(\operatorname{Var}(\hat{\beta})) \\
			&=&0+\sigma_Y^2 \operatorname{Tr}((X'X)^{-1})
	\end{eqnarray}	 
	and when correlations between covariates are strong, the matrix to invert is ill-conditioned, the variance explodes and so does the MSE.
	Another problem is that matricial inversion requires $n\geq p$. When it is not the case, a frequently used alternative is the Moore-Penrose generalized inverse \cite{katsikis2008fast}. In the followings, matrices inverses when $n<p$ will refer to this generalized inverse. Thus OLS can obtain some results even with $n<p$ (see section \ref{sectionsimul} ).
\subsection{Modelisation of the correlations}
Let now $X \in \mathcal{R}^{n\times p}$ be a set of $p$ correlated covariates.
We propose to explicitly define a family of $p_2$ internal regressions between covariates whith $I_2$ the set of indices of endogenous variables in $X$ (explained ones) and $I_1=\{I_1^1,\dots,I_1^p \}$ the set of the sets of indices of exogenous covariates (explaining ones) with $\forall j \notin I_2, I_1^j=\emptyset$. Then we have an explicit structure $S=(I_1,I_2,p_1,p_2)$ where $p_1=(p_1^1,\dots,p_1^{p_2})$ is the vector of the number of covariates in each internal regression. Thus we have $p_2=| I_2|$ and $p_1^j=|I_1^j|$ where $|.|$ represents the cardinal of an ensemble.

@@

The family of $p_2$ regressions within correlated covariates in $X$ is noted:
	\begin{equation}
		\forall j \in I_2: X^j_{|X_1,S}=X_1\delta_j+\varepsilon_j \textrm{ with } \varepsilon_j \sim(0,\sigma^2_jI_n) \label{SR}
	\end{equation}
		where $\delta_j \in \mathcal{R}^{(p-p_2)}$ are the vectors of the regression coefficients between the covariates (containing some zeros according to $I_1^j$).
It remains $p-p_2$ independent exogenous variables $X_1$: $\forall j \notin I_2 : X^j \sim f(\theta_j)$

@@

This generative model is conditional to $S$, the discrete structure model that is identifiable because we can't permute some regressions in (\ref{SR}) and obtain the same joint distribution $P(X,Y)$, the residuals ($\varepsilon_j$) would not stay Gaussian (see the appendices). These residuals are in the following supposed independent but one can suppose dependencies between them and then use appropriate tools to estimate them and the $\delta_j$ like SUR with Feasible Generalized Least Squares (FGLS) by Zellner \cite{SURzellner} or SPRING \cite{chiquetconf} as mentioned in the introduction.
\\

@@

\bibliography{biblio}{}
\bibliographystyle{plain}
\section{Appendices}
	\subsection{Identifiability of the structure}
	The model presented above relies on a discrete structure $S$ between the covariates. But to find it we need identifiability property to insure the MCMC will asymptotically find the true model. Identifiability of the discrete structure is asked in following terms: Is it possible to find another structure $\tilde{S}$ of linear regression between the covariates leading to the same joint distribution and marginal distributions? The answer is no. Thus $S$ is identifiable.

	If there are exact regressions ($\sigma^2_j=0$) in (\ref{SR}), the structure won't be identifiable. But it only means that several structure will have the same likelihood and they will have the same interpretation. So it's not really a problem. Moreover, when an exact sub-regression is found, we can delete one of the implied variables without any loss of information and the structure will define a list of variable from which to delete. %\textsc{CorReg} (Our R package) prints a warning to point out exact regressions when found.
	In the followings we suppose $\sigma^2_j\neq 0$.
